---
title: "Reproducible Research: Peer Assessment 1"


output: 
  html_document:
    keep_md: true
---

## Author: Camillo Falcone

## Loading and preprocessing the data
```{r read}
library(dplyr)
setwd("~/Desktop/Reproducible research/Assign_1")
data<-read.table("activity.csv", header = TRUE, sep = ",")
days <- group_by(data, date)
intervals <- group_by(data, interval)
total_by_day <- summarize(days, s1 = sum(steps, na.rm = TRUE))
avg_by_interval <- summarize(intervals, avg = mean(steps, na.rm = TRUE))
```


## What is mean total number of steps taken per day?

### Histogram

```{r histogram}
barplot(total_by_day$s1, main = "Total number of steps taken each day", ylab = "Number of steps")

```

### Mean and median
```{r mean}
m1 <- mean(total_by_day$s1)
m2 <- median(total_by_day$s1)
```
The mean value of the total number of steps taken per day is `r m1` while the median value is `r m2`.

### Time series plots

```{r time_series}

plot(avg ~ interval, avg_by_interval, type = "l")
```

## What is the average daily activity pattern?

### interval with the max avg of steps

```{r max}
max<- which(avg_by_interval$avg == max(avg_by_interval$avg))
```
The 5-minute interval, on average across all the days in the dataset, which contains the maximum number of steps is `r max`.

## Imputing missing values

```{r missing}
missing <- sum(is.na(data$steps))
library (data.table)
m<-data.table(merge (x = data, y = avg_by_interval, all.x = TRUE))
m<- m[order(m$date)]
m$avg <- as.integer(m$avg)
m[is.na(steps), steps := avg]
days_no_na <- group_by(m, date)
total_by_day_no_na <- summarize(days_no_na, s1 = sum(steps, na.rm = TRUE))
barplot(total_by_day_no_na$s1, main = "Total number of steps by days (with no missing values)", ylab = "Number of steps")
m3 <- mean(total_by_day_no_na$s1)
m4 <- median(total_by_day_no_na$s1)
```

The total nuber of missing values in the dataset is `r missing`.
By adopting the strategy of using the mean for the 5-minute interval when a NA value is found, the mean of the total number of steps taken per day is `r m3` while the median is `r m4`

## Are there differences in activity patterns between weekdays and weekends?

```{r weekends, fig.width=8, fig.height=8}
m<- m[,wd:=weekdays(as.Date(m$date))] 
levels(m$wd) <- c(levels(m$wd), c("weekend", "weekday")) 
m$wd[(m$wd == "Domenica") | (m$wd == "Sabato")] <- "weekend" 
m$wd[m$wd != "weekend"] <- "weekday" 

wd_intervals <- group_by(subset(m, wd == "weekday"), interval)
we_intervals <- group_by (subset(m, wd == "weekend"), interval )
wd_avg_by_interval <- summarize(wd_intervals, avg = mean(steps, na.rm = TRUE))
we_avg_by_interval <- summarize(we_intervals, avg = mean(steps, na.rm = TRUE))
par(mfcol= c(2,1))
plot(avg ~ interval, wd_avg_by_interval, type = "l", main = "Averaged across all working days")
plot(avg ~ interval, we_avg_by_interval, type = "l", main = "Averaged across all weekend days")


```
